Scalar Feedback based Joint Time-Frequency Precoder interpolation for MIMO-OFDM Systems

### 2020-01-09 paper/main.tex
Minor fixes
Add changes

For the adaptive channel tracking scheme, the arrows in
Fig.~\ref{fig:adpm-fig} show how previous quantized values along with
the 1-bit enhancement are used to find the new quantized value. Since
we use a single bit adaptation, the amount of feedback required is 1
bit for each scalar parameter that is being tracked. On average, this
can be reduced by half if we transmit only $\bf{\Theta}$ values for
one-time instance and only $\bf{\Phi}$ values for the next time
instant in an alternating fashion (the untransmitted value is assumed
to be the same as that at the previous instant). In this case, the
average number of bits required to quantize a channel matrix will be
$N_{T}(2N_{T} - N_{R})/2$. While tracking the $\phi_i$ parameters, is
must be noted they may change abruptly between $-\pi$ and $\pi$ due to
jumps of $2\pi$. This is avoided by unwrapping the phases to
facilitate continuous tracking (i.e., jumps in phase of
magnitude close to $2\pi$ have been eliminated).
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$, in keeping with the same notation as the previous
section. To construct the rest of the subcarriers, we will use
precoder parameters from both the past and future and time instances
to do the interpolation (using backtracking to ensure causality). In
Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
neighbouring subcarriers using bilinear interpolation (equivalent to
linear interpolation in two dimensions). These interpolated values and
then used to reconstruct the precoding matrices all together. Thus,
the approach here is able to exploit the time and frequency
correlation jointly to enhance performance.

@@

\section{Simulation and discussion}
\label{section3}
We now present simulations of the scalar feedback based joint
time-frequency predictive quantization outlined in the previous
section. In particular, we analyze the performance of the proposed
quantization and interpolation method for time-varying MIMO-OFDM
systems. The channel is simulated using the typical Jakes Model for
for the COST207 channel power delay
profiles~\cite{molisch2006cost259,cost1989cost}. The BER achieved by
the proposed predictive quantization based precoder is compared
against the ideal precoder for 100 channel evolutions in each run and
averaged over 10 channel realizations. Simulations are performed for
normalised Doppler values of $f_dT_s = 3.41\times 10^{-6}$ (corresponding
to a velocity of 37.2 km/h for $f_c = 1$ GHz) with 1024 subcarriers, and with
$f_dT_s = 1.56 \times 10^{-5}$ (velocity of 172.8 km/h) with 64
subcarriers. The sampling time period $T_s$ was $5\times10^{-8}$
s. For both the situations, we considered $N_T=4$ transmit antennas
and and $N_R=2$ receive antennas. Therefore, the total number of
parameters that determine the precoder is
$N_{R}(2N_{T} - N_{R}) = 12$, with five $\theta_l$s and and seven
$\phi_i$s. We assume that the spacing between fed back subcarrier
indices $p$ to be $33$ and $9$ for the slow and fast channel profiles
respectively (i.e. for the slower channels with 1024 subcarriers, the
subcarriers indices fed back would be $0, 33, 66, \ldots 1023$ and for
the faster channel, with 64 subcarriers, $0, 9, 18, \ldots 63$ would
be fed back).

For the first and the second simulation time instances, since the
precoder is not available at the transmitter, we use 2 bits for
initializing each parameter for effective representation of the
precoder. Therefore, the number of bits for each subcarrier =
$10\times 2$ = 20 (we note that $\phi_0 = \phi_4 = 0$ due to the
non-uniqueness of the SVD). Quantization is performed as given in
Section~\ref{section2}. We note that this is not significant, since,
amortized over long durations, this number becomes small. Subsequent
quantization takes place in the time-domain using one bit for each
parameter. Since the bit budget is small, if the channel varies
sufficiently slowly, either the $\theta_l$ or the $\phi_i$ parameters
are fed back alternately, while retaining the unsent parameters as
being to their previous values, as shown in
Fig.~\ref{fig:adpm-fig}.
% Here, $\bf{\Theta}$ is the collection of all
% $\theta_l$s and $\bf{\Phi}$ is the collection of all $\phi_i$s.
This brings down the bit budget to an average of 5 bits(the length of $\bf{\Theta}$)
and 7 bits (the length of $\bf{\Phi}$) over alternate feedback instances,
thereby resulting in an average feedback budget of 6 bits.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.78\columnwidth]{images/1024final_withtime}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 3.41\times 10^{-6}$, corresponding to a velocity of 37.8 km/h.}
\label{fig:ber_ped}
\end{center}
\end{figure}

@@

Fig.~\ref{fig:ber_veh} shows a similar comparison for a situation with
a higher speed. The faster variation of the channel taps with time
necessitates the use of a smaller number of subcarriers (viz. 64
subcarriers), thereby permitting more accurate tracking of precoders
on a per subcarrier basis. As in the previous case, we observe that
the scalar parameter-based approach outperforms manifold based
quantization by about 1 dB for higher SNRs. Naturally, the rate of
channel variation is quite fast, which limits the efficacy of the
proposed method in tracking the channel parameters.
\begin{figure}
\begin{center}
\includegraphics[width=0.78\columnwidth]{images/64final_withtime}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 1.56\times 10^{-5}$, corresponding to a velocity of 172.8 km/h.}
\label{fig:ber_veh}
\end{center}
\end{figure}

>>>>

For the adaptive channel tracking scheme, the arrows in
Fig.~\ref{fig:adpm-fig} show how previous quantized values along with
the 1-bit enhancement are used to find the new quantized value. Since
we use a single bit adaptation, the amount of feedback required is 1
bit for each scalar parameter that is being tracked. On average, this
can be reduced by half if we transmit only $\bf{\Theta}$ values for
one-time instance and only $\bf{\Phi}$ values for the next time
instant in an alternating fashion (the untransmitted value is assumed
to be the same as that at the previous instant). In this case, the
average number of bits required to quantize a channel matrix will be
$(N_{R}(2N_{T} - 1) - N_R^2)/2$ (from~\cite{4114278}). While tracking
the $\phi_i$ parameters, is must be noted they may change abruptly
between $-\pi$ and $\pi$ due to jumps of $2\pi$. This is avoided by
unwrapping the phases to facilitate continuous tracking (i.e., jumps
in phase of magnitude close to $2\pi$ have been eliminated).
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$, in keeping with the same notation as the previous
section. To construct the rest of the subcarriers, we will use
precoder parameters from both the past and future and time instances
to do the interpolation (using backtracking to ensure causality). In
Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
neighbouring subcarriers using bilinear interpolation (equivalent to
linear interpolation in two dimensions). These interpolated values and
then used to reconstruct the precoding matrices all together. Thus,
the approach here is able to exploit the time and frequency
correlation jointly to enhance performance.

@@

\section{Simulation and discussion}
\label{section3}
We now present simulations of the scalar feedback based joint
time-frequency predictive quantization outlined in the previous
section. In particular, we analyze the performance of the proposed
quantization and interpolation method for time-varying MIMO-OFDM
systems. The channel is simulated using the typical Jakes Model for
for the COST207 channel power delay
profiles~\cite{molisch2006cost259,cost1989cost}. The BER achieved by
the proposed predictive quantization based precoder is compared
against the ideal precoder for 100 channel evolutions in each run and
averaged over 10 channel realizations. Simulations are performed for
normalised Doppler values of $f_dT_s = 3.41\times 10^{-6}$ (corresponding
to a velocity of 37.2 km/h for $f_c = 1$ GHz) with 1024 subcarriers, and with
$f_dT_s = 1.56 \times 10^{-5}$ (velocity of 172.8 km/h) with 64
subcarriers. The sampling time period $T_s$ was $5\times10^{-8}$
s. For both the situations, we considered $N_T=4$ transmit antennas
and and $N_R=2$ receive antennas. Therefore, using the approach
outlined in~\cite{4114278}, the parameters that determine the precoder is
$N_{R}(2N_{T} - 1) - N_R^2 = 10$, with five $\theta_l$s and and seven
$\phi_i$s. We assume that the spacing between fed back subcarrier
indices $p$ to be $33$ and $9$ for the slow and fast channel profiles
respectively (i.e. for the slower channels with 1024 subcarriers, the
subcarriers indices fed back would be $0, 33, 66, \ldots 1023$ and for
the faster channel, with 64 subcarriers, $0, 9, 18, \ldots 63$ would
be fed back).

For the first and the second simulation time instances, since the
precoder is not available at the transmitter, we use 2 bits for
initializing each parameter for effective representation of the
precoder. Therefore, the number of bits for each subcarrier =
$10\times 2$ = 20 (we note that $\phi_0 = \phi_4 = 0$ due to the
non-uniqueness of the SVD). Quantization is performed as given in
Section~\ref{section2}. We note that this is not significant, since,
amortized over long durations, this number becomes small. Subsequent
quantization takes place in the time-domain using one bit for each
parameter. Since the bit budget is small, if the channel varies
sufficiently slowly, either the $\theta_l$ or the $\phi_i$ parameters
are fed back alternately, while retaining the unsent parameters as
being to their previous values, as shown in
Fig.~\ref{fig:adpm-fig}.
% Here, $\bf{\Theta}$ is the collection of all
% $\theta_l$s and $\bf{\Phi}$ is the collection of all $\phi_i$s.
This brings down the bit budget to an average of 5 bits (the length of
$\bf{\Theta}$) and 5 bits (the number of non-zero elements in
$\bf{\Phi}$) over alternate feedback instances, thereby resulting in
an average feedback budget of 5 bits.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.72\columnwidth]{images/1024final_withtime}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 3.41\times 10^{-6}$, corresponding to a velocity of 37.8 km/h.}
\label{fig:ber_ped}
\end{center}
\end{figure}

@@

Fig.~\ref{fig:ber_veh} shows a similar comparison for a situation with
a higher speed. The faster variation of the channel taps with time
necessitates the use of a smaller number of subcarriers (viz. 64
subcarriers), thereby permitting more accurate tracking of precoders
on a per subcarrier basis. As in the previous case, we observe that
the scalar parameter-based approach outperforms manifold based
quantization by about 1 dB for higher SNRs. Since, on a per subcarrier
basis, the channel variation is slower, the precoder tracking is
better than the previous case. Naturally, the rate of channel
variation is quite fast, which limits the efficacy of the proposed
method in tracking the channel parameters.
\begin{figure}
\begin{center}
\includegraphics[width=0.72\columnwidth]{images/64final_withtime}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 1.56\times 10^{-5}$, corresponding to a velocity of 172.8 km/h.}
\label{fig:ber_veh}
\end{center}
\end{figure}
